---
title: "Building gtfs and tidygtfs objects"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load test GTFS

```{r}
library(tidyverse)
library(tidytransit)

local_gtfs_path <- system.file("extdata", "google_transit_nyc_subway.zip", package = "tidytransit")
gtfs <- read_gtfs(local_gtfs_path)
names(gtfs)

```

## Explore GTFS tables.

The `.` sublist is for storing additional tables (additional to what? anything not strictly required?)

```{r}
gtfs$.
```

```{r}
gtfs$routes %>% class()
class(gtfs)
```

These are standard tibbles, not data.table as used by gtfsio for reading. The gtfs object has class `tidygtfs` as well as inheriting from `gtfs`.

## Try creating a new tidygtfs object

```{r}
txt_files <- list(
  agency = gtfs$agency,
  stops = gtfs$stops,
  routes = gtfs$routes,
  trips = gtfs$trips,
  stop_times = gtfs$stop_times,
  calendar = gtfs$calendar
)

new_gtfs <- gtfsio::new_gtfs(txt_files)
class(new_gtfs)

```

This is *just* a `gtfs` object, without the `tidygtfs` class. But it still validates okay:

```{r}
validate_gtfs(new_gtfs)
```

```{r}
read_gtfs
```

It seems `tidytransit` doesn't do too much more than gtfsio: just converts to tibble, convert dates and times, adds validation results and an empty `.` slot. However, those conversions rely on the table objects inheriting from `data.table`.

Let's try this on our `gtfs` object.

```{r}
g  <- gtfs
validation_result <- validate_gtfs(g)
g$. <- list()
# g <- tidytransit:::convert_times_to_hms(g)
# g <- tidytransit:::convert_dates(g)
# g <- tidytransit:::set_dates_services(g)
# g[names(g) != "."] <- lapply(g[names(g) != "."], 
#     dplyr::as_tibble)
# g <- gtfsio::new_gtfs(g)
# class(g) <- c("tidygtfs", "gtfs")
# attributes(g)$validation_result <- validation_result
# g
```

This fails at `tidytransit:::convert_times_to_hms(g) : inherits(gtfs_obj$stop_times, "data.table") is not TRUE`. Let's try converting to `data.table` first.

```{r}
g  <- gtfs
validation_result <- validate_gtfs(g)
g$. <- list()
g[names(g) != "."] <- lapply(g[names(g) != "."],
    data.table::as.data.table)
g <- tidytransit:::convert_times_to_hms(g)
g <- tidytransit:::convert_dates(g)
g <- tidytransit:::set_dates_services(g)
g[names(g) != "."] <- lapply(g[names(g) != "."],
    dplyr::as_tibble)
g <- gtfsio::new_gtfs(g)
class(g) <- c("tidygtfs", "gtfs")
attributes(g)$validation_result <- validation_result
class(g)
```

This seems to work! Let's make it into a function.

```{r}
make_tidygtfs <- function(gtfs_list) {
  g <- gtfsio::new_gtfs(gtfs_list)
  validation_result <- validate_gtfs(g)
  g$. <- list()
  g[names(g) != "."] <- lapply(g[names(g) != "."],
      data.table::as.data.table)
  g <- tidytransit:::convert_times_to_hms(g)
  g <- tidytransit:::convert_dates(g)
  g <- tidytransit:::set_dates_services(g)
  g[names(g) != "."] <- lapply(g[names(g) != "."],
      dplyr::as_tibble)
  g <- gtfsio::new_gtfs(g)
  class(g) <- c("tidygtfs", "gtfs")
  attributes(g)$validation_result <- validation_result
  g
}

test_txt_files <- list(
  agency = gtfs$agency,
  stops = gtfs$stops,
  routes = gtfs$routes,
  trips = gtfs$trips,
  stop_times = gtfs$stop_times,
  calendar = gtfs$calendar
)

test_gtfs <- make_tidygtfs(test_txt_files)
class(test_gtfs)

```


## Exporting as valid GTFS files

It could be worth doing this to create valid `tidygtfs` objects if we want to use any of the `tidytransit` functions for analysis, but it might be too much for just exporting to GTFS files for OTP use. In that case it might be easier just to adapt a simplified version of the code from `gtfsio::export_gtfs`.

Just as well `tidygtfs` writing seems to work okay:

```{r}
# gtfsio::export_gtfs
write_gtfs(test_gtfs, "test_gtfs_export.zip")
gtfs_from_new_file <- read_gtfs("test_gtfs_export.zip")
names(gtfs_from_new_file)
attributes(gtfs_from_new_file)$validation_result
```


### Important notes:

* ~~Dates in GTFS files need to be written as `yyyymmdd` text: beware of databases converting these to Date object-style integers.~~
* ~~GTFS times are written as `hh:mm:ss` text: beware of conversions to `hms` objects.~~
* We probably also want shapes to be included in a gtfs database.